Higher-Order unification for free!

### 2024-04-30 paper/main.tex
fix code snippet

In this paper we did study a compiler and simulation loop on a minimal language.
The actual implementation uses the Coq-Elpi meta language to both compile
the sequence of problems (the rules) and execute them, that is Elpi
plays the role of \Ho{} as well.

The main difference is that we cannot implement \hrun since it comes as part of
the runtime of the programming language. In particular the runtime iterates
\Ue, but \hstep also needs to check for \progress. Luckily Elpi extends~\cite{TASSI_2019}
$\lambda$Prolog with syntactic constraints (suspended goals) and
Constraint Handling Rules (CHR)~\cite{fruehwirth2017constraint} to operate on them.
An important design choice is that constraints are suspended on a user
computed list of unification variables, and are resumed as soon as one of them
is assigned, before any other existing goal is considered. In turn this
property enables a timely activation of links that grants~\fullref{prop:fidelity}.
For brevity we only provide two pseaudo-code snippets. The first one depicts
how \linketa are suspended or make \progress.

\begin{elpicode}
link-eta L R :- var L, maybe-eta R, !,
  declare_constraint (link-eta L R) [L,R].
link-eta L R :- not (var L), !, eta-progress-lhs L R.
link-eta L R :- not (maybe-eta R), !, contract R S, L = S.
\end{elpicode}

>>>>

In this paper we did study a compiler and a simulation loop on a minimal language.
The actual implementation uses the Coq-Elpi meta language to both compile
the sequence of problems (the rules) and execute them, that is Elpi
plays the role of \Ho{} as well.

The main difference is that we cannot implement \hrun since it comes as part of
the runtime of the programming language. In particular the runtime iterates
\Ue, but \hstep also needs to check links for \progress. Luckily Elpi extends
$\lambda$Prolog with constraints (suspended goals) and
Constraint Handling Rules (CHR) to operate on them~\cite{TASSI_2019,fruehwirth2017constraint}.
An important design choice is that constraints are suspended on a user
computed list of unification variables, and are resumed \emph{as soon as one of them
is assigned}, before any other existing goal is considered. In turn this
property allows for a timely activation of links that grants~\fullref{prop:fidelity}.
For brevity we only provide two pseaudo-code snippets. The first one depicts
how \linketa are suspended or make \progress.

\begin{elpicode}
link-eta L R :- var L, maybe-eta R, !,
  declare_constraint (link-eta L R) [L,R].
link-eta L R :- not (var L), !, eta-progress-lhs L R.
link-eta L R :- not (maybe-eta R), !, eta-progress-rhs L R.
\end{elpicode}
